---
title: "Multiple Linear Regression on Customer Data"
author : "Insofe Lab Session"
date: "July 02 , 2017"
output:
  html_document:
    toc_depth: 3
    toc_float: true
  

---
```{r}
# Go through the data and understand the attributes to get the info of the variables.
```

Write a code to clear the environment if needed
```{r}
rm(list=ls(all=(TRUE)))
```

Write a code to set the working directory.
```{r }
setwd("/Users/shwetabhat/Desktop/DataScience/CustomerData_Assignment")
```
Think and load the libraries as and when required in this place only as best practice
```{r}

```
Write a code to read the csv file as "data" as per required
```{r eval=TRUE }
customer_data=read.csv("/Users/shwetabhat/Desktop/DataScience/CustomerData_Assignment/CustomerData_Assignment.csv",header = T)
```
Write a code to know the names of the attributes
```{r}
names(customer_data)
colnames(customer_data)
```
Write a code to find the dimensions of the data
```{r}
dim(customer_data)
```

Write a code to see the head  and tail of the dataset atleast 10rows
```{r }
tail(customer_data,n = 10)
head(customer_data,n=10)
```
Write a code to see the data types of the attributes 
```{r }
str(customer_data)
```
Write a code to remove the unuseful variables and store the data 
```{r}
#Consider
city_data=customer_data$City
customer_data=within(customer_data,rm('City','CustomerID'))
##cust_data=customer_data[,-which(names(customer_data) == "City")]
```
Write a code to observe the variables and convert them into the required formats
```{r}
cat_attr=c("FavoriteGame","FavoriteChannelOfTransaction")
num_attr=c(setdiff(names(customer_data),cat_attr))
customer_data[num_attr]=data.frame(apply(customer_data[num_attr],2, function(x)as.numeric(x)))
#or customer_data[num_attr]=data.frame(lapply(customer_data[num_attr],2,FUN = as.numeric))
```
Check the changes again using str() command
```{r }
str(customer_data)
```
Write a code to get the numerical attributes out and store them as num_df and seperate categorical attributes.
```{r }
cat_df=customer_data[,cat_attr]
num_df=subset(customer_data,select=num_attr)
```
Write a Code to seperate the target
```{r}
target_attr=customer_data["TotalRevenueGenerated"]
num_df$TotalRevenueGenerated=NULL
```
Write a code for Normalizing the num_df
```{r}
library(vegan)
str(num_df)
num_df=decostand(num_df,method = "standardize")
```
Write a code to observe the correlation between the  attributes 
```{r}
cor(num_df)
```
Write a code to plot the corrplot of correlation between the attributes
```{r fig.height= 8, fig.width = 9}
library(corrplot)
corrplot(cor(num_df),method = "number")
```
Write a code to combine the numerical,categorical data along with the target 
```{r}
str(customer_data)
str(target_attr)
str(cat_df)
str(num_df)
combined_data=cbind(target_attr,num_df,cat_df)
```
Write a code to set the seed and comment why it is used.
```{r}
set.seed(29)
#set the seed so that the random number generated should be same every time you divide the data into train and test.
```
Write a code to get the train rows using sample 
```{r}
train_rows=sample(x=1:nrow(combined_data),size = 0.7*nrow(combined_data))
```
Write a code to get the train and test 
```{r}
train_data=combined_data[train_rows,]
test_data=combined_data[-train_rows,]
```

Write a code to just plot the graphs between attributes and targets
```{r}
par(mfrow=c(3,3))
plot(combined_data$NoOfChildren,combined_data$TotalRevenueGenerated,xlab="No Of Childern",ylab = "Total Revenue")
plot(combined_data$MinAgeOfChild,combined_data$TotalRevenueGenerated,xlab="MinAgeOfChild",ylab = "Total Revenue")
plot(combined_data$MaxAgeOfChild,combined_data$TotalRevenueGenerated,xlab="MaxAgeOfChild",ylab = "Total Revenue")
plot(combined_data$Tenure,combined_data$TotalRevenueGenerated,xlab="Tenure",ylab = "Total Revenue")
plot(combined_data$FrquncyOfPurchase,combined_data$TotalRevenueGenerated,xlab="FrquncyOfPurchase",ylab = "Total Revenue")
plot(combined_data$NoOfUnitsPurchased,combined_data$TotalRevenueGenerated,xlab="NoOfUnitsPurchased",ylab = "Total Revenue")
plot(combined_data$FrequencyOFPlay,combined_data$TotalRevenueGenerated,xlab="FrequencyOFPlay",ylab = "Total Revenue")
plot(combined_data$NoOfGamesPlayed,combined_data$TotalRevenueGenerated,xlab="NoOfGamesPlayed",ylab = "Total Revenue")
plot(combined_data$NoOfGamesBought,combined_data$TotalRevenueGenerated,xlab="NoOfGamesBought",ylab = "Total Revenue")

```
Write a code to form a linear regression model 
```{r}
linReg_model=lm(formula=TotalRevenueGenerated~NoOfChildren+MinAgeOfChild+MaxAgeOfChild+Tenure+FrquncyOfPurchase+NoOfUnitsPurchased+FrequencyOFPlay+NoOfGamesPlayed+NoOfGamesBought,data=train_data)
```
Write a code to plot the model
```{r}
par(mfrow = c(2,2))
plot(linReg_model)
```
Write a code to check the summary of the model 
```{r}
summary(linReg_model)
```
Write a code to predict the value of the target on the linear model 
```{r}
attr_without_target=num_attr[-length(num_attr)]
pred_values=predict(linReg_model,test_data[attr_without_target])
```
Write a code to use stepAIC
```{r}
library(MASS)
aic_model=stepAIC(linReg_model,direction = 'both')
```
Write a code to predict it on stepAIC
```{r}
predict_aic_values=predict(aic_model,test_data[attr_without_target])
```
Write a code to check the multicollinearity in the lm model 
```{r}
library(car)
vif(linReg_model)
```
Write a code to check the plots of the models
```{r}
par(mfrow=c(2,2))
plot(aic_model)
```

Write a code to check the multicollinearity problem
```{r}
vif(aic_model)
```
### Error Metrics for Regression

```{r}
#Mean Absolute Error (MAE)
mae <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error))
  
}

#Mean Squared Error (MSE)

mse <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(error^2)
  
}

#Root Mean Squared Error (RMSE)
rmse <- function(actual, predicted){
  
  error <- actual - predicted
  
  sqrt(mean(error^2))
  
}

#Mean Absolute Percentage Error (MAPE)
mape <- function(actual, predicted){
  
  error <- actual - predicted
  
  mean(abs(error/actual))*100
  
}

```
Write a code to evaluate the error in the prediction of AIC

```{r}
mae(test_data$TotalRevenueGenerated, predict_aic_values)
mse(test_data$TotalRevenueGenerated, predict_aic_values)
rmse(test_data$TotalRevenueGenerated, predict_aic_values)
mape(test_data$TotalRevenueGenerated, predict_aic_values)

# OR

library(DMwR)

regr.eval(test_data$TotalRevenueGenerated, predict_aic_values)

```
Write a code to evaluate the error on the prediction of Linear Regression
```{r}
mae(test_data$TotalRevenueGenerated, pred_values)
mse(test_data$TotalRevenueGenerated, pred_values)
rmse(test_data$TotalRevenueGenerated, pred_values)
mape(test_data$TotalRevenueGenerated, pred_values)

# OR

regr.eval(test_data$TotalRevenueGenerated, pred_values)

```

#As stepAIC removed Tenure because high colinearity is invovled.

```{r}
improved_ln_model=lm(formula=TotalRevenueGenerated~NoOfChildren+MinAgeOfChild+MaxAgeOfChild+FrquncyOfPurchase+NoOfUnitsPurchased+FrequencyOFPlay+NoOfGamesPlayed,data=train_data)

#check the model summary.
summary(improved_ln_model)

pred_values_improved=predict(improved_ln_model,test_data[attr_without_target])
vif(improved_ln_model)
regr.eval(test_data$TotalRevenueGenerated, pred_values_improved)

```
