---
title: "Association Rules "
author: "Sudharani Rangapale"
date: "July 21, 2017"
output: html_document
---
# Problem 3 - Assignment
### Association Rules for "Flight Delay" dataset: Generate the rules and identify the patterns. 

1. Read the data into R

```{r, echo=T}
FlightDelays=read.csv("/Users/shwetabhat/Desktop/DataScience/A_Rules_Assignment/FlightDelays.csv")
```


2. Look at the summary of all the variables and convert the following variables as factors 
        + Weather 
        + DAY_WEEK 
        + Flight Status
```{r}
str(FlightDelays)
  summary(FlightDelays)
  FlightDelays$Weather=as.factor(as.character((FlightDelays$Weather)))
  FlightDelays$DAY_WEEK =as.factor(as.character((FlightDelays$DAY_WEEK )))
  FlightDelays$Flight.Status=as.factor(as.character((FlightDelays$Flight.Status)))
  
```

3. Bin the numeric variable 'CRS_DEP_TIME' into 4 bins as follows: If time is between 6 AM to 12 noon code as 1 and If the time is between 2 , etc.
```{r}

FlightDelays$CRS_DEP_TIME = factor(cut(FlightDelays$CRS_DEP_TIME, breaks = c(599, 1200, 1800, 2400, Inf), labels = c("1", "2", "3", "4")))
```

4. Convert the data frame in a transactions object. Look at the first 6 transactions to understand how 'arules' library requires data to be transformed. Use the below R code to view the transactions data 
```{r}
library(arules)
Flightdelays_transc=as(FlightDelays,"transactions")
inspect(Flightdelays_transc)



```

5. Apply 'arules' algorithm and play with various support, lift and confidence values.
```{r}

rules=apriori(Flightdelays_transc,parameter=list(sup=0.2,conf=0.6,target="rules"))
```

6. Inspect all the rules.
```{r}
inspect(rules)

```

7. Filter the rules with specific LHS and RHS conditions 
E.g.; Filter the rules with Flighstatus=0
```{r}
rules.sorted <- sort(rules, by="lift")
inspect(rules.sorted)

```

8. Filter redundant rules if any
```{r}
subset.matrix <- is.subset(rules.sorted, rules.sorted,sparse = FALSE)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1
which(redundant)
rules.pruned <- rules.sorted[!redundant]
inspect(rules.pruned)
```

9. Pot and visualize the rules
```{r}
library(arulesViz)
plot(rules.pruned)
```