---
title: "ARules_Assignment"
author: "Shweta"
date: "8/5/2017"
output: html_document
---

# Problem 3 - Assignment
### Association Rules for "Flight Delay" dataset: Generate the rules and identify the patterns. 

1. Read the data into R

```{r, echo=T}
rm(list = ls(all=T))

flight_data<-read.csv(file = "/Users/shwetabhat/Desktop/DataScience/A_Rules_Assignment/FlightDelays.csv", header = T, sep = ",")

sum(is.na((flight_data)))
```


2. Look at the summary of all the variables and convert the following variables as factors 
        + Weather 
        + DAY_WEEK 
        + Flight Status
```{r}
str(flight_data)

flight_data$Weather<-as.factor(as.character(flight_data$Weather))
flight_data$DAY_WEEK<-as.factor(as.character(flight_data$DAY_WEEK))
flight_data$Flight.Status<-as.factor(as.character(flight_data$Flight.Status))

str(flight_data)
```

3. Bin the numeric variable 'CRS_DEP_TIME' into 4 bins as follows: If time is between 6 AM to 12 noon code as 1 and If the time is between 2 , etc.
```{r}
##000 to 559 will not have any values

flight_data$CRS_DEP_TIME = factor(cut(flight_data$CRS_DEP_TIME, breaks = c(0,599, 1200, 1800, 2400), labels = c(1, 2, 3, 4)))


unique(flight_data$CRS_DEP_TIME)

table(flight_data$CRS_DEP_TIME)

summary(flight_data)

```

4. Convert the data frame in a transactions object. Look at the first 6 transactions to understand how 'arules' library requires data to be transformed. Use the below R code to view the transactions data 
```{r}
library(arules)

##Convert into Transactional objects
flight_trans <-  as(flight_data, "transactions")

summary(flight_trans)

itemInfo(flight_trans)

#inspect(flight_trans)
itemFrequencyPlot(flight_trans)
```

5. Apply 'arules' algorithm and play with various support, lift and confidence values.
```{r}
rules1<-apriori(flight_trans, parameter = list(support = 0.1, conf = 0.1, target = 'rules', minlen = 3))
rules1

inspect(rules1[1:10])
rules2<-apriori(flight_trans, parameter = list(support = 0.1, conf = 0.3, target = 'rules', minlen = 3))
rules2
```

6. Inspect all the rules.
```{r}
rules1=sort(rules1, by = c("support", "confidence"))

rules2=sort(rules2, by = c("support", "confidence"))
inspect(rules2[1:15])
inspect(rules1[1:15])
```

7. Filter the rules with specific LHS and RHS conditions 
E.g.; Filter the rules with Flighstatus=0
```{r}
## filtering rhs(rules) get flightstatus 
#minlen=3 corresponds to rules with atleast 3 elements.
filtered_rule1 <- subset(rules1, subset = rhs %in% 
                        "Flight.Status=0")#has only rules object.
filtered_rule1_df <- as(subset(rules1, subset = rhs %in% 
                        "Flight.Status=0"),"data.frame")#rules converted into dataframe

filtered_rule2_df <- as(subset(rules1, subset = rhs %in% 
                        "Flight.Status=1"),"data.frame")
filtered_rule2 <- subset(rules1, subset = rhs %in% 
                        "Flight.Status=1")
```

8. Filter redundant rules if any
```{r}

#Check for count of redundant rules.
table(duplicated(filtered_rule1))
table(is.redundant(filtered_rule1))
table(duplicated(filtered_rule2))
table(is.redundant(filtered_rule2))
#Check for redundancy in rules.
which(is.redundant(filtered_rule1))

# Rules Pruning.

subset.matrix <- is.subset(filtered_rule1, filtered_rule1,sparse = FALSE)
dim(subset.matrix)
subset.matrix[lower.tri(subset.matrix, diag=T)] <- NA
redundant <- colSums(subset.matrix, na.rm=T) >= 1
which(redundant)

#remove the redundancy.
rules.pruned <- filtered_rule1[!redundant]
inspect(rules.pruned[1:15])
```

9. Pot and visualize the rules
```{r}
library(arulesViz)
plot(rules.pruned)
```