---
title: "Association Rules "
author: " "
date: "July 21, 2017"
output: html_document
---
# Problem 3 - Assignment
### Association Rules for "Flight Delay" dataset: Generate the rules and identify the patterns. 

1. Read the data into R

```{r, echo=T}
rm(list = ls(all=T))

setwd("C:\\Users\\C5215696\\Desktop\\Data Science\\Association-Rules")

flight_data<-read.csv(file = "FlightDelays.csv", header = T, sep = ",")

sum(is.na((flight_data)))

colSums(is.na(flight_data))
```


2. Look at the summary of all the variables and convert the following variables as factors 
        + Weather 
        + DAY_WEEK 
        + Flight Status
```{r}
str(flight_data)

flight_data$Weather<-as.factor(flight_data$Weather)
flight_data$DAY_WEEK<-as.factor(flight_data$DAY_WEEK)
flight_data$Flight.Status<-as.factor(flight_data$Flight.Status)

str(flight_data)
```

3. Bin the numeric variable 'CRS_DEP_TIME' into 4 bins as follows: If time is between 6 AM to 12 noon code as 1 and If the time is between 2 , etc.
```{r}
##As data have starting value from 6 AM, we will have only 3 bins. 000 to 559 will not have any levels.
flight_data$CRS_DEP_TIME<-cut(flight_data$CRS_DEP_TIME, breaks = c(1159,1759,2359, 559), labels = c(1,2,3))

unique(flight_data$CRS_DEP_TIME)

table(flight_data$CRS_DEP_TIME)

summary(flight_data)

head(flight_data)

dim(flight_data)

```

4. Convert the data frame in a transactions object. Look at the first 6 transactions to understand how 'arules' library requires data to be transformed. Use the below R code to view the transactions data 
```{r}
library(arules)

##Transactional objects
flight_trans <-  as(flight_data, "transactions")

summary(flight_trans)

itemInfo(flight_trans)

inspect(flight_trans[2000:2010])

image(flight_trans)
```

5. Apply 'arules' algorithm and play with various support, lift and confidence values.
```{r}
rules1<-apriori(flight_trans, parameter = list(support = 0.1, conf = 0.1, target = 'rules', minlen = 3))

rules1

inspect(rules1[1:10])

head(quality(rules1))

rules2<-apriori(flight_trans, parameter = list(support = 0.1, conf = 0.2, target = 'rules', minlen = 3))

head(quality(rules2))

rules2

rules3<-apriori(flight_trans, parameter = list(support = 0.1, conf = 0.5, target = 'rules', minlen = 3))

rules3
```

6. Inspect all the rules.
```{r}

inspect(sort(rules1[1:20], by = "lift"))
inspect(sort(rules1[1:20], by = c("support", "confidence")))

inspect(sort(rules2[1:20], by = "lift"))
inspect(sort(rules2[1:20], by = c("support", "confidence")))

inspect(sort(rules3[1:20], by = "lift"))
inspect(sort(rules3[1:20], by = c("support", "confidence")))
```

7. Filter the rules with specific LHS and RHS conditions 
E.g.; Filter the rules with Flighstatus=0
```{r}
## filtering only rhs(rules) as we need to find which rules affect flightstatus the most and more over rules with > 2 elements will be rendered because we supplied minlen=3 
rules.filter1 <- sort(subset(rules1, subset = rhs %in% 
                        c("Flight.Status=0", "Flight.Status=1")), by=c("support", "confidence"))

rules.filter2 <- sort(subset(rules2, subset = rhs %in% 
                        c("Flight.Status=0", "Flight.Status=1")), by=c("support", "confidence"))

rules.filter3 <- sort(subset(rules3, subset = rhs %in% 
                        c("Flight.Status=0", "Flight.Status=1")), by=c("support", "confidence"))

```

8. Filter redundant rules if any
```{r}

rules.filter1
rules.filter2
rules.filter3

itemFrequency(rules.filter1[01:57]@lhs, type = "absolute")
itemFrequency(rules.filter2[01:57]@lhs, type = "absolute")
itemFrequency(rules.filter3[01:56]@lhs, type = "absolute")

table(duplicated(rules.filter1))
table(is.redundant(rules.filter1))
table(duplicated(rules.filter2))
table(is.redundant(rules.filter2))
table(duplicated(rules.filter3))
table(is.redundant(rules.filter3))

# After applyig filters for necessary condition(flightstatus), we get same rules in all three variables. Hence using rule.filter1 for our analysis.

which(is.redundant(rules.filter1))

#Removing redudant filters from rule.filter1
rules1_upd<-rules.filter1[!is.redundant(rules.filter1)]
typeof(rules1_upd)
table(is.redundant(rules1_upd))
```

9. Pot and visualize the rules
```{r}
library(arulesViz)
plot(rules1_upd)
```