---
title: "DTAssignmet"
author: "Shweta"
date: "8/5/2017"
output: html_document
---

# C5.0 Trees

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}
rm(list = ls(all=TRUE))
```

## Goal

* The goal of this activity is to predict wether a patient has liver disease or not based on various patient related attributes


## Agenda 

* Get the data

* Data Pre-processing

* Build a model

* Predictions

* Communication


## Reading & Understanding the Data

### Read the Data

Make sure the dataset is located in your current working directory, or else you can change your working directory using the "setwd()" function.

```{r}
diseaseInfo_data=read.csv("/Users/shwetabhat/Desktop/DataScience/DecisionTrees_assignment/ilpd_data.csv",header = T)
```

### Understand the data

* Use the str(), summary(), head() and tail() functions to get the dimensions and types of attributes in the dataset

* The dataset has 582 observations and 11 variables

- __The variable descriptions are given below:__

1 - age	 : Age of the patient

2 - gender : Gender of the patient

3 - TB : Total Bilirubin content

4 - DB : Direct Bilirubin content

5 - alk_phos : Alkaline Phosphotase content

6 - alamine : Alamine Aminotransferase content

7 - aspartate : Aspartate Aminotransferase content

8 - TP : Total Protiens content

9 - albumin : 	Albumin content

10 - A/G : Ratio of Albumin and Globulin 

11 - Disease : Whether the patient has liver disease or not 


```{r}
str(diseaseInfo_data)
summary(diseaseInfo_data)

head(diseaseInfo_data)
tail(diseaseInfo_data)
```


## Data Pre-processing

### Verify Data Integrity

* Verify if the dataset has missing values

```{r}
colSums(is.na(diseaseInfo_data))

# Impute missing values using knnImputation() function in both the train and test datasets
library(DMwR)
diseaseInfo_data = knnImputation(diseaseInfo_data)
```

* Verify the data types assigned to the variables in the dataset

```{r}
str(diseaseInfo_data)
```

### Split the Data into train and test sets

* Use stratified sampling to split the data into train/test sets (70/30)

* Use the createDataPartition() function from the caret package to do stratified sampling

```{r}
library(caret)
set.seed(111)
trainIndex <- createDataPartition(diseaseInfo_data$disease, p = .7, list = F)

train_data <- diseaseInfo_data[trainIndex, ]
test_data <- diseaseInfo_data[-trainIndex, ]
```

## Build a  Decision Tree

### Model the tree

* Use Quinlan's C5.0 decision tree algorithm implementation from the C50 package to build your decision tree

```{r}
library(C50)
c5_tree=C5.0(disease ~ .,train_data)
```

* Build a rules based tree

```{r}
c5_rules <- C5.0(disease ~ . , train_data, rules = T)
```

### Variable Importance in trees

* Find the importance of each variable in the dataset

```{r}
C5imp(c5_tree, metric = "usage")
```

### Rules from trees

* Understand the summary of the returned c5.0 rules based on the decision tree model


```{r}
summary(c5_rules)
```


### Plotting the tree

* Call the plot function on the tree object to visualize the tree

```{r, fig.width= 35, fig.height=15}
plot(c5_tree)
```


## Evaluating the model

### Predictions on the test data

* Evaluate the decision tree using the standard error metrics on test data

```{r}
predicted <- predict(c5_tree, test_data)
```

* Report error metrics for classification on test data

```{r}
library(caret)
ilpd=confusionMatrix(predicted, test_data$disease)
```


# CART Trees

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}

rm(list=ls(all=TRUE))

```

* The classification and regression trees use gini index in place of the gain ratio (based on information gain) used by the ID3 based algorithms, such as c4.5 and c5.0

## Goal

* The goal of this activity is to predict the heating load of a residential building, if the building parameters are given

* Hence, in the future architects would be able to build more energy efficient buildings as they can optimize the building parameters to reduce the heating load

## Agenda 

* Get the data

* Data Pre-processing

* Build a model

* Predictions

* Communication


## Reading & Understanding the Data

### Read the Data

* Make sure the dataset is located in your current working directory, or else you can change your working directory using the "setwd()" function.

```{r}

building_energy_data=read.csv("/Users/shwetabhat/Desktop/DataScience/DecisionTrees_assignment/building_energy.csv",header = T)

```

### Understand the data

* Use the str(), summary(), head() and tail() functions to get the dimensions and types of attributes in the dataset

* The dataset has 768 observations and 9 variables

```{r}
str(building_energy_data)
summary(building_energy_data)
head(building_energy_data)
tail(building_energy_data)

```

* The variable names are self explanatory, for further information visit http://www.sciencedirect.com/science/article/pii/S037877881200151X

## Data Pre-processing

### Verify Data Integrity

* Verify if the dataset has missing values

```{r}
sum(is.na(building_energy_data))
```

* Verify the data types assigned to the variables in the dataset

```{r}
str(building_energy_data)
```

### Split the Data

* Split the data into train/test sets (70/30)

```{r}
set.seed(200)

train_rows <- sample(1:nrow(building_energy_data), 0.7*nrow(building_energy_data))

train_data <- building_energy_data[train_rows, ]

test_data <- building_energy_data[-train_rows, ]
```

## Build a Regression Tree

### Model the tree

* Use the rpart package to build a cart tree to predict the heating load

```{r}

library(rpart)
reg_tree <- rpart(heating_load ~ ., train_data)
printcp(reg_tree)
```

### Tree Explicability

* Print the variable importance

```{r}
reg_tree$variable.importance
```

* Plot the regression tree

```{r, fig.width=8, fig.height=5}
library(rpart.plot)
rpart.plot(reg_tree)
```

## Evaluation on Test Data

* Report error metrics on the test data

```{r}
preds_reg <- predict(reg_tree, test_data)
```

```{r}
library(DMwR)
regr.eval(test_data$heating_load, preds_reg)
```