---
title: "SVMandRandomFoestAssignment"
author: "Shweta"
date: "8/28/2017"
output: html_document
---
  
###Clear the Global Environment
```{r}
rm(list=ls(all=TRUE))
```
#Lodaing Libraries
```{r}
library(DMwR)
library(dummies)
library(e1071)
```

# Reading data.
```{r}
sar_data = read.csv(file = "/Users/shwetabhat/Downloads/assignment-svm-random-forest-ShwetaGBhat-master/SaratogaHouses.csv", sep = ",")
```

# Looking at the data 
```{r}
print(str(sar_data))
print(summary(sar_data))
```

#Checking for missing values.
```{r}
sum(is.na(sar_data))
```

#Creating dummies for SVM as SVM runs only on Numerical data.
```{r}
str(sar_data)
dummified_Val = data.frame(dummy(sar_data$heating))
sar_data[,"heating"] = NULL
sar_data = cbind(sar_data,dummified_Val)

dummified_Val = data.frame(dummy(sar_data$fuel))
sar_data[,"fuel"] = NULL
sar_data = cbind(sar_data,dummified_Val)

dummified_Val = data.frame(dummy(sar_data$sewer))
sar_data = cbind(sar_data,dummified_Val)
sar_data[,"sewer"] = NULL

dummified_Val = data.frame(dummy(sar_data$waterfront))
sar_data = cbind(sar_data,dummified_Val)
sar_data[,"waterfront"] = NULL

dummified_Val = data.frame(dummy(sar_data$newConstruction))
sar_data = cbind(sar_data,dummified_Val)
sar_data[,"newConstruction"] = NULL

dummified_Val = data.frame(dummy(sar_data$centralAir))
sar_data = cbind(sar_data,dummified_Val)
sar_data[,"centralAir"] = NULL
```
#Splitting the data into train and Test
```{r}
train_rows = sample(1:nrow(sar_data), nrow(sar_data)*.7)
trainData = sar_data[train_rows,]
testData = sar_data[-train_rows,]
```
#Building SVM Regression
```{r}
svmModel = svm(x = trainData[,2:ncol(trainData)], y = trainData[,1], kernel = "linear")
summary(svmModel)
```
#Model Evaluation on SVM
```{r}
regr.eval(trainData$price, predict(svmModel, trainData[,2:ncol(trainData)]))
regr.eval(testData$price, predict(svmModel, testData[,2:ncol(testData)]))
```

#Building Random Forest
# Clear all environment variables
```{r}
rm(list = ls(all = TRUE))
```
#Lodaing Libraries
```{r}
library(DMwR)
library(randomForest)
```
# Reading the data
```{r}
sar_data = read.csv(file = "/Users/shwetabhat/Downloads/assignment-svm-random-forest-ShwetaGBhat-master/SaratogaHouses.csv", sep = ",")
```
#Data observation.
```{r}
print(str(sar_data))
print(summary(sar_data))
sum(is.na(sar_data))
```
#Splitting the data into train and Test
```{r}
train_rows = sample(1:nrow(sar_data), nrow(sar_data)*.7)
trainData = sar_data[train_rows,]
testData = sar_data[-train_rows,]
```
#Building Random Forest 
```{r}
rmModel = randomForest(price ~ ., trainData,ntree=1200)
summary(rmModel)
```
#Evaluate RM Model.
```{r}
regr.eval(trainData$price, predict(rmModel, trainData[,2:ncol(trainData)]))
regr.eval(testData$price, predict(rmModel, testData[,2:ncol(testData)]))
```
